---
title: "Subjective and autonomic nervous system responses to affective pictures in people with frequent nightmares"
output:
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
if (!require(tidyverse)) install.packages("tidyverse")
if (!require(lme4)) install.packages("lme4")
if (!require(lmerTest)) install.packages("lmerTest")
if (!require(patchwork)) install.packages("patchwork")

library(tidyverse)
library(fs) # Part of extended tidyverse
library(lme4)
library(lmerTest)
library(patchwork)
```

# Aims

Theories suggest that people with frequent nightmares may have disturbed emotional processing, however this claim has not been supported by experimental evidence. This research investigates the following claims: 
People with frequent nightmares...  
A) ...have a heightened reactivity to disturbing stimuli  
B) ...have a slower habituation to disturbing stimuli

# Methods
Data was collected from participants with and without frequent nightmares. After a first accommodating night, participants were exposed to neutral and negative IAPS pictures in three blocks, while skin conductance and electrocardiogram was recorded. Arousal and valence of the pictures were evaluated after exposure. Subsequently, participant spent a night at the sleep lab, and took part in a recall session in the morning when where New pictures were shown along with the original ones.

# Data analysis strategy
- We use linear mixed-effects modeling using the lme4 package
- The random structure is determined by the Zuur (2009) method to determine the random and fixed structure
- Outcome variables are **subjective arousal, valence, skin conductance response (SCR), and heart rate deceleration (HRD)** (see processing details for each outcome measure below)
- We used **group** (control/nightmare) **block** (1/2/3) **picture** (neutral/negative) variables as fixed effects
- We start with a full model, and remove effects sequentially that does not contribute to the model
- Model comparisons are based on BIC: difference that are larger than 2 are considered important.

```{r data read, echo=FALSE, message=FALSE, warning=FALSE}
# Participant data
participants <- read_csv("analysis_data/participants.csv")

# Marker data
markers <- 
    read_csv("analysis_data/markers.csv") %>% 
    # Remove flanker pictures
    filter((session == "emot" & order >= 4 & order <= 63) |
            session == "recall" & order >= 4 & order <= 43) %>% 
    # Set factor levels for picture
    mutate(picture = as_factor(picture, levels = c("Neutral", "Negative")))
    
markers_wo_time <- select(markers, -time)

# Physiological data
hr_data <- read_csv("all_data/hr/hr_linear.csv") # This is "raw" HR data
scr_data <- read_csv("analysis_data/scr_data.csv")
hrd_data <- read_csv("analysis_data/hrd_data.csv")

# Extraction picture sets from marker data
pic_sets <- 
    markers_wo_time %>% 
    drop_na(block) %>% 
    filter(session == "emot") %>%
    select(id, stimulus) %>% 
    inner_join(participants %>% select(id, pic_set), by = "id") %>% 
    distinct(stimulus, pic_set) %>% 
    arrange(pic_set)

set_a <- pic_sets %>% filter(pic_set == "A") %>% pull(stimulus)
set_b <- pic_sets %>% filter(pic_set == "B") %>% pull(stimulus)

```

## Participant info (TBA)
- Table about the average age, gender by group
- Differences between groups in gender, age, and psych scales
- Put subjective, scr, hrd data into one dataset (?)

## Prepare data
```{r prepare scr data}
all_data <-
    participants %>% 
    left_join(scr_data, by = "id") %>% 
    left_join(hrd_data, by = c("id", "session", "block", "stimulus")) %>% 
    left_join(markers_wo_time, by = c("id", "session", "stimulus", "block")) %>% 
    # Remove outliers larger that 3.5 sds
    filter(abs(scr_std) < 3.5 & abs(hrd_std) < 3.5) %>%
    # Add whether a picture in the recall session is New or Old, a block is Pre-sleep or Post-sleep, and set factor levels
    mutate(novelty = case_when(
               session == "recall" & pic_set == "A" & stimulus %in% set_a ~ "Old",
               session == "recall" & pic_set == "A" & !(stimulus %in% set_a) ~ "New",
               session == "recall" & pic_set == "B" & stimulus %in% set_b ~ "Old",
               session == "recall" & pic_set == "B" & !(stimulus %in% set_b) ~ "New",
               TRUE ~ NA_character_) %>% 
                   factor(., levels = c("Old", "New")),
            sleep = case_when(
                block == 3 ~ "Pre-sleep",
                session == "recall" & novelty == "Old" ~ "Post-sleep",
                TRUE ~ NA_character_) %>% 
                   factor(levels = c("Pre-sleep", "Post-sleep")),
            block = if_else(session == "recall", NA_integer_, block),
            picture = factor(picture, levels = c("Neutral", "Negative"))
    )

# Create a df where data is aggregated for participants by session, block, group, etc.
df_sum <-
    all_data %>% 
    group_by(id, session, block, group, picture, novelty, sleep) %>% 
    summarise(scr = mean(scr_sqrt, na.rm = TRUE),
              hrd = mean(hrd, na.rm = TRUE),
              arousal = mean(arousal, na.rm = TRUE),
              valence = mean(valence, na.rm = TRUE)) %>% 
    ungroup()

# This dataset is needed for the model that contains sleeping
sleep_df <-
    df_sum %>% 
    drop_na(sleep) %>% 
    select(-session, -block, -novelty)

# Dataset for plotting the sleep data
sleep_plot_df <-
    all_data %>% 
    drop_na(sleep) %>% 
    group_by(group, picture, sleep) %>% 
    summarise(scr_mean = mean(scr_sqrt, na.rm = TRUE),
              scr_se = sd(scr_sqrt, na.rm = TRUE)/sqrt(n()),
              hrd_mean = mean(hrd, na.rm = TRUE),
              hrd_se = sd(hrd, na.rm = TRUE)/sqrt(n()),
              arousal_mean = mean(arousal, na.rm = TRUE),
              arousal_se = sd(arousal, na.rm = TRUE)/sqrt(n()),
              valence_mean = mean(valence, na.rm = TRUE),
              valence_se = sd(valence, na.rm = TRUE)/sqrt(n())) %>% 
    ungroup()

skimr::skim(df_sum)
```

# Subjective arousal

```{r plot subjective arousal, echo=FALSE}
# Subjective arousal
plot_emot_arousal <-
    all_data %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(arousal, na.rm = TRUE),
              Se = sd(arousal, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(1, 9)) +
        scale_x_continuous("Block", breaks = 1:3) +
        scale_y_continuous("Mean (SEM) of subjective arousal") +
        guides(linetype = "none") +
        facet_grid(group ~ session) +
        theme(strip.text.y = element_blank())

plot_recall_arousal <-
    all_data %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(arousal, na.rm = TRUE),
              Se = sd(arousal, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(1, 9)) +
        scale_x_discrete("Novelty") +
        scale_y_continuous(NULL) +
        facet_grid(group ~ session) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank()) +
    labs(linetype = "Picture")

plot_emot_arousal + 
    plot_recall_arousal +
    plot_layout(widths = c(3,2))

```

## Emotion habituation model (arousal)
```{r modeling arousal}
arousal_lm <- lm(arousal ~ group * block * picture, data = df_sum)
arousal_i <- lmer(arousal ~ group * block * picture + (1 | id), data = df_sum)
arousal_is <- lmer(arousal ~ group * block * picture + (group : picture | id), data = df_sum)

# Remove 3 way interaction
arousal_best <- arousal_i
# Note that we use the random intercept model, however the random intercept+slope model may have a better fit (although it fails to converge properly). The optimal random structure is yet to be found.
arousal_3 <- update(arousal_best, . ~ . - group : block : picture)

arousal_best <- arousal_3

# Remove 2-way interactions
arousal_2_1 <- update(arousal_best, . ~ . - group : block) # Drop
arousal_2_2 <- update(arousal_best, . ~ . - group : picture) # Drop
arousal_2_3 <- update(arousal_best, . ~ . - block : picture) # Drop

arousal_best <- update(arousal_best, . ~ . - group:block - group:picture - block:picture)

# We keep all main effects
summary(arousal_best)
```

- There is a clear and strong effect of the picture type as negative pictures are associated with a higher subjective arousal compared to the neutral pictures. 

## Sleep habituation model (arousal)
```{r arousal sleep habituation}
sleep_plot_df %>% 
    ggplot() +
    aes(x = sleep, y = arousal_mean, ymin = arousal_mean - arousal_se, ymax = arousal_mean + arousal_se, group = picture) +
    geom_point() +
    geom_line(aes(linetype = picture), size = 1.2) +
    geom_errorbar(width = .2) +
    facet_grid(group ~ .)

lmer(arousal ~ group * picture * sleep + (1 | id), data = sleep_df) %>% 
    summary()
```

# Subjective valence

```{r plot subjective valence, echo=FALSE}
# Subjective valence
plot_emot_valence <-
    all_data %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(valence, na.rm = TRUE),
              Se = sd(valence, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(1, 9)) +
        scale_x_continuous("Block", breaks = 1:3) +
        scale_y_continuous("Mean (SEM) of subjective valence") +
        guides(linetype = "none") +
        facet_grid(group ~ session) +
        theme(strip.text.y = element_blank())

plot_recall_valence <-
    all_data %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                                   session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(valence, na.rm = TRUE),
              Se = sd(valence, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(1, 9)) +
        scale_x_discrete("Novelty") +
        scale_y_continuous(NULL) +
        facet_grid(group ~ session) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank()) +
        labs(linetype = "Picture")

plot_emot_valence + 
    plot_recall_valence +
    plot_layout(widths = c(3,2))

```
## Emotion habituation model (valence)
```{r modeling valence}
valence_lm <- lm(valence ~ group * block * picture, data = df_sum)
valence_i <- lmer(valence ~ group * block * picture + (1 | id), data = df_sum)
valence_is <- lmer(valence ~ group * block * picture + (block : picture | id), data = df_sum)

# Remove 3 way interaction
valence_best <- valence_i
# Note that we use the random intercept model, however the random intercept+slope model may have a better fit (although it fails to converge properly). The optimal random structure is yet to be found.
valence_3 <- update(valence_best, . ~ . - group : block : picture)

valence_best <- valence_3

# Remove 2-way interactions
valence_2_1 <- update(valence_best, . ~ . - group : block) # Drop
valence_2_2 <- update(valence_best, . ~ . - group : picture) # Drop
valence_2_3 <- update(valence_best, . ~ . - block : picture) # Keep

valence_best <- update(valence_best, . ~ . - group:block - group:picture)

# We keep all main effects
summary(valence_best)
```

- There is a clear and strong effect of the picture type as negative pictures are associated with a lower subjective valence compared to the neutral pictures. 
## Sleep habituation model (valence)
```{r valence sleep habituation}
sleep_plot_df %>% 
    ggplot() +
    aes(x = sleep, y = valence_mean, ymin = valence_mean - valence_se, ymax = valence_mean + valence_se, group = picture) +
    geom_point() +
    geom_line(aes(linetype = picture), size = 1.2) +
    geom_errorbar(width = .2) +
    facet_grid(group ~ .)

lmer(valence ~ group * picture * sleep + (1 | id), data = sleep_df) %>% 
    summary()
```

# Skin Conductance Response (SCR)
## Data processing
- Skin conductance was recorded using a sampling rate of 32 Hz
- Ledalab toolbox was used to down-sample data to 16 Hz, use continuous decomposition analysis with optimized parameters to separate tonic (SCL) and phasic (SCR) skin conductance. SCR was calculated using a response window of 1 to 4 sec after stimuli presentation, with a minimum amplitude threshOld criterion of 0.01 muS. 
- SCR was square root transformed and scaled for each participant

## Exploratory data analysis on data quality
### Number of markers/participants in the data analysis
Most data can be matched to markers. However ...
- all recordings are missing for 41, 43, and 54.

```{r scr markers}
participants %>% 
    left_join(scr_data, by = "id") %>% 
    arrange(id, session) %>% 
    mutate(id = as_factor(id), 
           session = as_factor(session)) %>% 
    group_by(id, session) %>% 
    count() %>% 
    mutate(n = if_else(n == 1, 0L, n)) %>% 
    ggplot() +
    aes(x = fct_rev(id), 
        y = n, 
        fill = session, 
        label = n) +
    geom_col(position = "dodge") +
    geom_text() +
    coord_flip() +
    labs(title = "Successfully matched markers for each session",
         y = "Number of matched markers", 
         x = "Participant id")
```

## Emotion habituation model (SCR)
```{r plotting SCR, echo=FALSE}
# Plotting SCR
plot_emot_scr <-
    all_data %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(scr_sqrt, na.rm = TRUE),
              Se = sd(scr_sqrt, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(-0.4, .75)) +
        scale_x_continuous("Block", breaks = 1:3) +
        scale_y_continuous("Mean (SEM) of skin counductance response") +
        guides(linetype = "none") +
        facet_grid(group ~ session) +
        theme(strip.text.y = element_blank())

plot_recall_scr <-
    all_data %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(scr_sqrt, na.rm = TRUE),
           Se = sd(scr_sqrt, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(-0.4, .75)) +
        scale_x_discrete("Novelty") +
        scale_y_continuous(NULL) +
        facet_grid(group ~ session) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank())

plot_emot_scr + 
    plot_recall_scr +
    plot_layout(widths = c(3,2))

```

```{r scr model building}
# Using Zuur method to select random structure
# Simple linear regression to justify lmer
model_lm <- lm(scr ~ group * block * picture, data = df_sum)
# Random intercept
model_i <- lmer(scr ~ group * block * picture + (1 | id), data = df_sum)
# Random intercept and splope
model_is <- lmer(scr ~ group * block * picture + (block : picture | id), data = df_sum)

# anova(model_i, model_lm)
# anova(model_i, model_is)

# The random intercept model is the best
#summary(model_i)

# Remove 3 way interaction
model_best <- model_i
model_3 <- update(model_best, . ~ . - group : block : picture)

# anova(model_best, model_3)
model_best <- model_3  # Drop 3-way interaction
#summary(model_best)

# Remove 2-way interactions
model_2_1 <- update(model_best, . ~ . - group : block)
model_2_2 <- update(model_best, . ~ . - group : picture)
model_2_3 <- update(model_best, . ~ . - block : picture)

# anova(model_best, model_2_1) # Drop
# anova(model_best, model_2_2) # Drop
# anova(model_best, model_2_3) # Keep
model_best <- update(model_best, . ~ . - group:block - group:picture)

# Remove main effects
# Remove 2-way interactions
model_1_1 <- update(model_best, . ~ . - block)
model_1_2 <- update(model_best, . ~ . - group)
model_1_3 <- update(model_best, . ~ . - picture)

# anova(model_best, model_1_1) # Can't remove
# anova(model_best, model_1_2) # Drop
# anova(model_best, model_1_3) # Keep

model_best <- update(model_best, . ~ . - group)

summary(model_best)
```

- There was no main or interaction effect of nightmare group membership on SCR 
- SCR decreased in each subsequent block, showing signs of habituation
- Negative pictures were associated with larger SCR

## Sleep habituation model (SCR)

```{r scr sleep habituation}
sleep_plot_df %>% 
    ggplot() +
    aes(x = sleep, y = scr_mean, ymin = scr_mean - scr_se, ymax = scr_mean + scr_se, group = picture) +
    geom_point() +
    geom_line(aes(linetype = picture), size = 1.2) +
    geom_errorbar(width = .2) +
    facet_grid(group ~ .)

lmer(scr ~ group * picture * sleep + (1 | id), data = sleep_df) %>% 
    summary()


```

Finding: There was no evidence for a difference in SCR before and after sleep.

# Heart rate deceleration (HRD)
## Data processing
- ECG data was collected using a sampling rate of 1024 Hz
- Movement drift was removed from the data using a moving average with a window of 1024 samples
- Data was down-sampled to 128 Hz
- IBIs were identified using a global threshOld in Artiifact, and corrected manually.
- Artifacts were identified using the Berntson method, and replaced using cubic spline interpolation
- Artifact corrected IBIs were used to estimate heart rate for each half second, using linear interpolation.
- Heart rate deceleration (HRD) was calculated as the minimum heart rate in a 3s window after presentation for each stimulus, relative to the moment of presentation
- HRD was negated, square root transformed and scaled for each participant
- HRD was averaged for each stimuli type (neutral vs. negative) for each participant, and each block.

## Exploratory data analysis on data quality
### Number of artifacts in the IBI data
First we check the percentage of artifacts in each recording. According to the following plot, the number of corrected artifacts in the sample was low (<3%).

```{r hrd artifacts}
ibi <-
tibble(artifact_file = dir_ls(path = "all_data/artifact_corrected_ibi_berntson/", regexp = "IBI_artifacts_IBIs_from_"),
       corrected_file = dir_ls(path = "all_data/artifact_corrected_ibi_berntson/", regexp = "IBI_artifactCorrected_IBIs_from_")
       ) %>% 
    transmute(id = str_extract(artifact_file, "\\d+"),
              session = str_replace(artifact_file, ".*_\\d+_(\\w+)_ECG.txt.csv$", "\\1"),
              artifact_data = map(artifact_file, ~read_csv(.x, col_types = "di")),
              corrected_data = map(corrected_file, ~read_csv(.x, col_types = "d")),
              data = map2(artifact_data,
                          corrected_data,
                          ~bind_cols(.x, .y))
                      ) %>%
    unnest(data) %>% 
    rename(ibi = IBI, ibi_corrected = IBI_artefactCorrected)


ibi %>% 
    group_by(id, session) %>% 
    summarise(artifact_n = sum(artifact),
              artifact_p = sum(artifact)/n()) %>% 
    ggplot() +
    aes(x = artifact_p, color = session, fill = session) +
    scale_x_continuous(labels = scales::percent_format()) +
    geom_freqpoly(size = 1.3, alpha = .7) +
    labs(title = "Frequency of artifacts as a percentage of the whole recoring",
         y = "Number of cases",
         x = "Percentage of the whole recording")
```

### Number of markers/participants in the data analysis
Most data can be matched to markers. However ...
- all recordings are missing for 26, 41, 43 and 54.
- the complete session of emot is missing for 04, 42; and recall is missing for 33, 34.
- participant 47 has only 6 successfully matched markers in the emot session.

```{r}
participants %>% 
    left_join(hrd_data, by = "id") %>% 
    mutate(id = as_factor(id), 
           session = as_factor(session)) %>% 
    group_by(id, session) %>% 
    count() %>% 
    mutate(n = if_else(n == 1, 0L, n)) %>% 
    ggplot() +
    aes(x = fct_rev(id), 
        y = n, 
        fill = session, 
        label = n) +
    geom_col(position = "dodge") +
    geom_text() +
    coord_flip() +
    labs(title = "Successfully matched markers for each session",
         y = "Number of matched markers",
         x = "Participant id")

```

```{r hrd momentary data preparation}
# Nudge data points to the closest half second. This decreases marker accuracy somewhat (from perfect accuracy, they become quarter-of-a-second accurate, which should be enough if we consider that HRD is calculated in a 3 second window)

markers_hrd <- 
    markers %>% 
    select(id, session, block, stimulus, time, picture) %>% 
    mutate(time = plyr::round_any(time,.5),
           block = if_else(session == "recall", NA_integer_, block)) %>% 
    arrange(id, session, time)

hr_6s <-
    hr_data %>% 
    # Join markers to the heart rate data
    left_join(markers_hrd, 
              by = c("id", "session", "time")) %>% 
    group_by(session, id) %>% 
    # Fill the missing data (last value carry forward)
    fill(block, stimulus, picture) %>%
    # select(-order, -valence, -arousal, -familiarity) %>% 
    group_by(id, session, stimulus, block) %>% 
    # Keep only the 6 seconds after stimuli presentation
    slice(0:12) %>% 
    # Calculate relative time and heart rate (compared to stimuli presentation)
    mutate(rel_time = time - first(time),
           hr_change = hr - first(hr)) %>% 
    ungroup() %>% 
    # Join only those that were not excluded from the sample 
    right_join(participants, by = "id") %>% 
    drop_na(session, block) %>% 
    arrange(id, session, block, stimulus, rel_time)
```

## Exploring HR reaction after stimuli presentation
```{r}
# Change in HR after stimuli presentation (all participants/sessions)
hr_6s %>% 
    group_by(picture, rel_time) %>% 
    summarise(hr_change = mean(hr_change, na.rm = TRUE)) %>% 
    ggplot() +
    aes(x = rel_time, y = hr_change, color = picture) +
    geom_smooth(method = "loess") +
    labs(title = "Change in heart rate after stimuli presentation (all participants/sessions)",
         x = "Time (s)",
         y = "Heart rate change (bpm)")
```

```{r}
# Change by session
hr_6s %>% 
    group_by(id, session, group, picture, rel_time) %>% 
    summarise(hr_change = mean(hr_change, na.rm = TRUE)) %>% 
    ggplot() +
        aes(x = rel_time, y = hr_change, color = picture) +
        geom_smooth(method = "loess") +
        facet_grid(group~session) +
        labs(title = "Change in heart rate after stimuli presentation in sessions",
             x = "Time (s)",
             y = "Heart rate change (bpm)")
```


```{r}
# Change in heart rate after stimuli presentation in the emot session by block
hr_6s %>% 
    filter(session == "emot") %>% 
    group_by(id, group, picture, rel_time, block) %>% 
    summarise(hr_change = mean(hr_change, na.rm = TRUE)) %>% 
    arrange(id, block, picture, rel_time) %>% # Just for checking values
    ggplot() +
        aes(x = rel_time, y = hr_change, color = picture) +
        geom_smooth(method = "loess") +
        facet_grid(group~block) +
        labs(title = "Change in heart rate after stimuli presentation in the emot session by block",
             x = "Time (s)",
             y = "Heart rate change (bpm)")
```


```{r}
# Change in heart rate after stimuli presentation in all sessions
hr_6s %>% 
    group_by(group, picture, rel_time) %>% 
    summarise(hr_change = mean(hr_change, na.rm = TRUE)) %>% 
    ggplot() +
    aes(x = rel_time, y = hr_change, color = picture) +
    geom_smooth(method = "loess") +
    facet_grid(group ~ .) +
    labs(title = "Change in heart rate after stimuli presentation in all sessions",
         x = "Time (s)",
         y = "Heart rate change (bpm)")
```

## Calculate Hear rate deceleration
The minimum of heart rate after stimuli presentation in a 3 second frame.
```{r}
# This block should be moved to a different script
hrd <-
    hr_data %>% 
    # Join markers to the heart rate data
    left_join(markers_hrd, 
              by = c("id", "session", "time")) %>% 
    group_by(session, id) %>% 
    # Fill the missing data (last value carry forward)
    fill(block, stimulus, picture) %>%
    group_by(id, session, block, stimulus) %>% 
    # Keep only the 6 seconds after stimuli presentation
    slice(0:6) %>% 
    # Calculate relative time and heart rate (compared to stimuli presentation)
    mutate(rel_time = time - first(time),
           hr_change = hr - first(hr)) %>% 
    group_by(id) %>% 
    mutate(hr_change_std = scale(hr_change)) %>% 
    ungroup() %>% 
    filter(abs(hr_change_std) < 3.5) %>% 
    # Join only those that were not excluded from the sample 
    right_join(participants, by = "id") %>% 
    drop_na(session, block) %>% 
    group_by(id, session, block, stimulus, picture, group, gender, pic_set) %>% 
    summarise(hrd = min(hr_change, na.rm = TRUE)) %>% 
    ungroup() %>% 
    mutate(novelty = case_when(
        session == "recall" & pic_set == "A" & stimulus %in% set_a ~ "Old",
        session == "recall" & pic_set == "A" & !(stimulus %in% set_a) ~ "New",
        session == "recall" & pic_set == "B" & stimulus %in% set_b ~ "Old",
        session == "recall" & pic_set == "B" & !(stimulus %in% set_b) ~ "New",
        TRUE ~ NA_character_) %>% 
            factor(., levels = c("Old", "New"))) %>% 
    mutate(block = case_when(session == "recall" & novelty == "New" ~ NA_integer_,
                             TRUE ~ block)) %>% 
    arrange(id, session, block, stimulus)

hrd %>% 
    ggplot() +
    aes(x = hrd) +
    geom_histogram()
```


```{r}
plot_emot_hrd <-
    all_data %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(hrd, na.rm = TRUE),
              Se = sd(hrd, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(-1.5, -4.5)) +
        scale_x_continuous("Block", breaks = 1:3) +
        scale_y_continuous("Mean (SEM) of Hear Rate Deceleration") +
        guides(linetype = "none") +
        facet_grid(group ~ session) +
        theme(strip.text.y = element_blank()) +
        NULL

plot_recall_hrd <-
    all_data %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(hrd, na.rm = TRUE),
              Se = sd(hrd, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(-1.5, -4.5)) +
        scale_x_discrete("Novelty") +
        scale_y_continuous(NULL) +
        facet_grid(group ~ session) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank()) +
        NULL

plot_emot_hrd + 
    plot_recall_hrd +
    plot_layout(widths = c(3,2))
```

## Emotion habituation model (HRD)
```{r hrd emotion habituation}
# Using Zuur method to select random structure
# Make the recall the 4th block
# hrd_emot <- 
#     mutate(df_sum, block = if_else(session == "recall", 4L, block))

# Linear regression
hrd_lm <- lm(hrd ~ group * block * picture, data = df_sum)
# Random intercept
hrd_i <- lmer(hrd ~ group * block * picture + (1 | id) , data = df_sum)
# Random intercept and splope
hrd_is <- lmer(hrd ~ group * block * picture + (picture * block | id), data = df_sum)

#anova(hrd_i, hrd_lm)
#anova(hrd_i, hrd_is)

# The random intercept model is the best
#summary(hrd_i)

# Remove 3 way interaction
hrd_best <- hrd_i
hrd_3 <- update(hrd_best, . ~ . - group : block : picture)

#anova(hrd_best, hrd_3)
hrd_best <- hrd_3  # Drop 3-way interaction
#summary(hrd_best)

# Remove 2-way interactions
hrd_2_1 <- update(hrd_best, . ~ . - group : block)
hrd_2_2 <- update(hrd_best, . ~ . - group : picture)
hrd_2_3 <- update(hrd_best, . ~ . - block : picture)

# anova(hrd_best, hrd_2_1) # Drop
# anova(hrd_best, hrd_2_2) # Keep
# anova(hrd_best, hrd_2_3) # Drop
hrd_best <- update(hrd_best, . ~ . - group:block - block : picture)

# summary(hrd_best)

# Remove main effects
# Remove 2-way interactions
hrd_1_1 <- update(hrd_best, . ~ . - block)
hrd_1_2 <- update(hrd_best, . ~ . - group)
hrd_1_3 <- update(hrd_best, . ~ . - picture)

# anova(hrd_best, hrd_1_1) # Can't remove
# anova(hrd_best, hrd_1_2) # Drop
# anova(hrd_best, hrd_1_3) # Keep

hrd_best <- hrd_1_2
summary(hrd_best)
```

## Sleep habituation model (HRD)
```{r hrd sleep habituation}
sleep_plot_df %>% 
    ggplot() +
    aes(x = sleep, y = hrd_mean, ymin = hrd_mean - hrd_se, ymax = hrd_mean + hrd_se, group = picture) +
    geom_point() +
    geom_line(aes(linetype = picture), size = 1.2) +
    geom_errorbar(width = .2) +
    facet_grid(. ~ group)

lmer(hrd ~ group * picture * sleep + (1 | id), data = sleep_df) %>% 
    summary()
```


# Cortisol
## Data processing
Salivary cortisol was collected at bedtime, and in the next morning at awakening,  15 minutes and 30 minutes after awakening on both days.
...details of cortisol measurement...

- Missing cortisol data (n = 3) was imputed using linear interpolation.
- AUCg and AUCi were calculated for each participant and each night, using the bedtime, awakening, awakening +15, and awakening +30 data points.

```{r}
cort_df_raw <-read_csv("analysis_data/cortisol_data.csv")

cort_df <- 
    participants %>%
    left_join(cort_df_raw, by = "id") %>%
    group_by(id, day) %>%
    # If the first or last cortisol value is missing, we have to drop it
    drop_na(cortisol) %>% 
    # Interpolating cortisol data
    mutate(cortisol_full = approx(x = cort_time,
                                  y = cortisol,
                                  xout = cort_time)$y) %>%
    ungroup() %>% 
    mutate(event =  fct_inorder(event))

cort_auc <-
    cort_df %>% 
    group_by(id, group, day) %>% 
    nest() %>% 
    mutate(auc = map(data, ~calculate_auc(.x, 
                                 time = "cort_time",
                                 measurement = "cortisol_full",
                                 type = "both"))) %>% 
    unnest(auc) %>% 
    select(-data)

```

## Exploratory data analysis on data quality
### Number of markers/participants in the data analysis

```{r}
participants %>% 
    left_join(cort_df_raw, by = "id") %>% 
    mutate(id = as_factor(id), 
           day = as_factor(day)) %>% 
    group_by(id, day) %>% 
    count(cort_missing = is.na(cortisol)) %>% 
    filter(!cort_missing) %>% 
    mutate(n = if_else(n == 1, 0L, n)) %>% 
    ggplot() +
        aes(x = fct_rev(id), 
            y = n, 
            fill = day, 
            label = n) +
        geom_col(position = "dodge") +
        geom_text() +
        coord_flip() +
        labs(title = "Successfully matched cortisol samples at each day",
             y = "Number of matched cortisol samples",
             x = "Participant id")
```

## Cortisol change over time

```{r}
cort_df %>% 
    group_by(group, day, cort_time) %>% 
    summarise(cortisol_mean = mean(cortisol, na.rm = TRUE),
              cortisol_se = sd(cortisol, na.rm = TRUE)/sqrt(n()),
              minute_mean = mean(cort_minute, na.rm = TRUE),
              minute_se = sd(cort_minute, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
    aes(x = minute_mean, 
        y = cortisol_mean, 
        ymin = cortisol_mean - cortisol_se,
        ymax = cortisol_mean + cortisol_se,
        xmin = minute_mean - minute_se,
        xmax = minute_mean + minute_se,
        group = group) +
    geom_line(aes(color = group), size = 1.2) +
    geom_point(aes(color = group), size = 2) +
    geom_linerange() +
    geom_errorbarh(height = 0) +
    facet_wrap(~day) +
    ggtitle("Cortisol change by day, time, and group with respect to absolute time")
```

```{r}
cort_df %>% 
    group_by(group, day, cort_time, event) %>% 
    summarise(cortisol_mean = mean(cortisol, na.rm = TRUE),
              cortisol_se = sd(cortisol, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
    aes(x = event, 
        y = cortisol_mean, 
        ymin = cortisol_mean - cortisol_se,
        ymax = cortisol_mean + cortisol_se,
        group = group) +
    geom_line(aes(color = group), size = 1.2) +
    geom_point(aes(color = group), size = 2) +
    geom_linerange() +
    facet_wrap(~day) +
    ggtitle("Cortisol change by day, time, and group with respect to time points")
```

## Difference in AUC
AUC was calculated with respect to ground (AUCg) and with respect to baseline (AUCi). Linear Mixed-effects models were fitted on the data, with the respective AUC metric as outcome, and group membership (control/nightmare) and day (first/second), and their interaction as predictors. 

### AUCg
There was no significant effect of the predictors on AUCg

```{r}
aucg_model <- lmer(auc_g ~ group * day + (1|id), data = cort_auc)
summary(aucg_model)
```

### AUCi
We found a significant interaction between group and day, and a main effect of day.
Closer inspection of the values suggest that the nightmare group showed a smaller decrease in AUCi from the first to the second day, compared to the control group.

that the nightmare group showed a lower AUCi decrease 

```{r}
auci_model <- lmer(auc_i ~ group * day + (1|id), data = cort_auc)
summary(auci_model)
```

```{r}
cort_auc %>% 
    group_by(group, day) %>% 
    summarise(AUCi_mean = mean(auc_i, na.rm = TRUE),
              AUCi_se = sd(auc_i, na.rm = TRUE)/sqrt(n()),
              AUCg_mean = mean(auc_g, na.rm = TRUE),
              AUCg_se = sd(auc_g, na.rm = TRUE)/sqrt(n())
            ) %>% 
    ungroup() %>% 
    gather(metric, value, -group, -day) %>% 
    separate(metric, c("type", "stat"), sep = "_") %>% 
    spread(stat, value) %>% 
    ggplot() +
        aes(x = day, 
            y = mean, 
            ymin = mean - se,
            ymax = mean + se,
            group = group) +
        geom_line(aes(color = group), size = 1.8) +
        geom_point(aes(color = group), size = 3) +
        geom_linerange() +
        facet_wrap(~type) +
        labs(title = "Changes in AUCg and AUCi on two nights",
             y = "Mean AUC (+-SE)",
             x = NULL)
```





