---
title: "Autonomic nervous system responses to affective pictures in people with frequent nightmares"
output: html_notebook
---

```{r setup, include=FALSE}
if (!require(tidyverse)) install.packages("tidyverse")
if (!require(lme4)) install.packages("lme4")
if (!require(lmerTest)) install.packages("lmerTest")
if (!require(patchwork)) install.packages("patchwork")

library(tidyverse)
library(fs) # Part of extended tidyverse
library(lme4)
library(lmerTest)
library(patchwork)
```

# Aims

Theories suggest that people with frequent nightmares may have disturbed emotional processing, however this claim has not been supported by experimental evidence. This research investigates the following claims: 
People with frequent nightmares ...
A) ... have a hightened reactivity to disturbing stimuli
B) ... have a slower habituation to disturbing stimuli

# Methods
Data was collected from participants with and without frequent nightmares. Participants were exposed to neutral and negative IAPS pictures, while skin conductance and electrocardiogram was recorded. 

## Data processing  
### Skin conductance Response (SCR)  
- Skin conductance was recorded using a sampling rate of 32 Hz
- Ledalab toolbox was used to downsample data to 16 Hz, use continuous decomposition analysis with optimized parameters to separate tonic (SCL) and phasinc (SCR) skin conductance. SCR was calculated using a response window of 1 to 4 sec after stimuli presentation, with a minimum amplitude threshold criterion of 0.01 muS. 

### Heart rate deceleration (HDR)
- ECG data was collected using a sampling rate of 1024 Hz
- Movement drift was removed from the data using a moving average with a window of 1024 samples
- Data was down-sampled to 128 Hz
- IBIs were identified using a global threshold in Artiifact, and corrected manually.
- (?) Artifacts were identified using the Berntson (?) method, and replaced using cubic spline interpolation
- Artifact corrected (?) IBIs   were used to estimate heart rate for each half second, using cubic spline interpolation.
- Heart rate deceleration (HRD) was calculate as the minimum heart rate in a 3s window after presentation for each stimulus, relative to the moment of presentation
- HRD was averaged for each stimuli type (neutral vs. negative) for each participant, and each block.

```{r data read, echo=FALSE, message=FALSE, warning=FALSE}
participants <- read_csv("analysis_data/participants.csv")
markers <- read_csv("analysis_data/markers.csv")
scr_data <- read_csv("analysis_data/scr_era.csv")

# Extraction picture sets from marker data
pic_sets <- 
    markers %>% 
    drop_na(block) %>% 
    filter(session == "emot") %>%
    select(id, stimulus) %>% 
    inner_join(participants %>% select(id, pic_set), by = "id") %>% 
    distinct(stimulus, pic_set) %>% 
    arrange(pic_set)

set_a <- pic_sets %>% filter(pic_set == "A") %>% pull(stimulus)
set_b <- pic_sets %>% filter(pic_set == "B") %>% pull(stimulus)

```


```{r prepare scr data}
scr_era <-
    scr_data %>% 
    group_by(id) %>% 
    # Standardize values, sqrt + standardize values
    mutate(scr_std = scale(scr),
           scr_sqrt = sqrt(scr) %>% scale()) %>% 
    ungroup() %>% 
    # Remove outliers larger that 3.5 sds
    filter(abs(scr_std) < 3.5) %>% 
    # Add whether a picture in the recall session is new or not
    mutate(novelty = case_when(
               session == "recall" & pic_set == "A" & stimulus %in% set_a ~ "old",
               session == "recall" & pic_set == "A" & !(stimulus %in% set_a) ~ "new",
               session == "recall" & pic_set == "B" & stimulus %in% set_b ~ "old",
               session == "recall" & pic_set == "B" & !(stimulus %in% set_b) ~ "new",
               TRUE ~ NA_character_) %>% 
               factor(., levels = c("old", "new"))) %>% 
    mutate(block = if_else(session == "recall", NA_integer_, block),
           picture = factor(picture, levels = c("Neutral", "Negative"))) 

scr_era %>% 
    ggplot() +
        aes(x = scr_sqrt) +
        geom_histogram()

```

```{r plotting SCR}
# Plotting SCR
plot_emot_scr <-
    scr_era %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(scr_sqrt, na.rm = TRUE),
              Se = sd(scr_sqrt, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(-0.4, .75)) +
        scale_x_continuous("Block", breaks = 1:3) +
        scale_y_continuous("Mean (SEM) of skin counductance response") +
        guides(linetype = "none") +
        facet_grid(group ~ session) +
        theme(strip.text.y = element_blank())

plot_recall_scr <-
    scr_era %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(scr_sqrt, na.rm = TRUE),
           Se = sd(scr_sqrt, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(-0.4, .75)) +
        scale_x_discrete("Novelty") +
        scale_y_continuous(NULL) +
        facet_grid(group ~ session) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank())

plot_emot_scr + 
    plot_recall_scr +
    plot_layout(widths = c(3,2))

```

```{r plot subjective arousal}
# Subjective arousal
plot_emot_arousal <-
    scr_era %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(arousal, na.rm = TRUE),
              Se = sd(arousal, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(1, 9)) +
        scale_x_continuous("Block", breaks = 1:3) +
        scale_y_continuous("Mean (SEM) of subjective arousal") +
        guides(linetype = "none") +
        facet_grid(group ~ session) +
        theme(strip.text.y = element_blank())

plot_recall_arousal <-
    scr_era %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(arousal, na.rm = TRUE),
              Se = sd(arousal, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
        aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
        geom_point() +
        geom_errorbar(width = .2) +
        geom_line(aes(linetype = picture), size = 1) +
        coord_cartesian(ylim = c(1, 9)) +
        scale_x_discrete("Novelty") +
        scale_y_continuous(NULL) +
        facet_grid(group ~ session) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank()) +
    labs(linetype = "Picture")

plot_emot_arousal + 
    plot_recall_arousal +
    plot_layout(widths = c(3,2))

```

```{r plot subjective valence}
# Subjective valence
plot_emot_valence <-
    scr_era %>%
    filter(session == "emot") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                               session == "recall" ~ "Recall")) %>% 
    group_by(session, block, group, picture) %>% 
    summarise(Mean = mean(valence, na.rm = TRUE),
              Se = sd(valence, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
    aes(x = block, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
    geom_point() +
    geom_errorbar(width = .2) +
    geom_line(aes(linetype = picture), size = 1) +
    coord_cartesian(ylim = c(1, 9)) +
    scale_x_continuous("Block", breaks = 1:3) +
    scale_y_continuous("Mean (SEM) of subjective valence") +
    guides(linetype = "none") +
    facet_grid(group ~ session) +
    theme(strip.text.y = element_blank())

plot_recall_valence <-
    scr_era %>%
    filter(session == "recall") %>% 
    mutate(session = case_when(session == "emot" ~ "Emotion induction",
                                   session == "recall" ~ "Recall")) %>% 
    group_by(session, group, picture, novelty) %>% 
    summarise(Mean = mean(valence, na.rm = TRUE),
              Se = sd(valence, na.rm = TRUE)/sqrt(n())) %>% 
    ggplot() +
    aes(x = novelty, y = Mean, group = picture, ymin = Mean - Se, ymax = Mean + Se) +
    geom_point() +
    geom_errorbar(width = .2) +
    geom_line(aes(linetype = picture), size = 1) +
    coord_cartesian(ylim = c(1, 9)) +
    scale_x_discrete("Novelty") +
    scale_y_continuous(NULL) +
    facet_grid(group ~ session) +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank()) +
    labs(linetype = "Picture")

plot_emot_valence + 
    plot_recall_valence +
    plot_layout(widths = c(3,2))

```

# Model building
```{r}
# Create a df where data is aggregated for participants by session, block, group, etc.
df_sum <-
    scr_era %>% 
    group_by(id, session, block, group, picture, novelty) %>% 
    summarise(scr = mean(scr_sqrt, na.rm = TRUE),
              arousal = mean(arousal, na.rm = TRUE),
              valence = mean(valence, na.rm = TRUE)) %>% 
    ungroup()

```

# Using Zuur method to select random structure
I've used the BIC value (>2) to decide which model had the best fit.

```{r model building}
# Simple linear regression to justify lmer
model_lm <- lm(scr ~ group * block * picture, data = df_sum)
# Random intercept
model_i <- lmer(scr ~ group * block * picture + (1 | id), data = df_sum)
# Random intercept and splope
model_is <- lmer(scr ~ group * block * picture + (block : group : picture | id), data = df_sum)

# anova(model_i, model_lm)
# anova(model_i, model_is)

# The random intercept model is the best
#summary(model_i)

# Remove 3 way interaction
model_best <- model_i
model_3 <- update(model_best, . ~ . - group : block : picture)

# anova(model_best, model_3)
model_best <- model_3  # Drop 3-way interaction
#summary(model_best)

# Remove 2-way interactions
model_2_1 <- update(model_best, . ~ . - group : block)
model_2_2 <- update(model_best, . ~ . - group : picture)
model_2_3 <- update(model_best, . ~ . - block : picture)

# anova(model_best, model_2_1) # Drop
# anova(model_best, model_2_2) # Drop
# anova(model_best, model_2_3) # Keep
model_best <- update(model_best, . ~ . - group:block - group:picture)

# Remove main effects
# Remove 2-way interactions
model_1_1 <- update(model_best, . ~ . - block)
model_1_2 <- update(model_best, . ~ . - group)
model_1_3 <- update(model_best, . ~ . - picture)

# anova(model_best, model_1_1) # Can't remove
# anova(model_best, model_1_2) # Drop
# anova(model_best, model_1_3) # Keep

model_best <- update(model_best, . ~ . - group)
```

## Findings
- There was no main or interaction effect of nightmare group membership on SCR 
- SCR decreased in each subsequet block
- Negative pictures were associated with larger SCR

### Final model
```{r final model}
summary(model_best)
```


# Habituation before and after sleep
Finding: Ther was no evidence for 
```{r}
sleep_df <-
    scr_era %>% 
    filter(session == "emot" & block == 3L |
               session == "recall" & novelty == "old") %>% 
    mutate(sleep = case_when(session == "emot" ~ "pre-sleep",
                             session == "recall" ~ "post-sleep") %>% 
               factor(., c("pre-sleep", "post-sleep"))) %>% 
    select(-session, -block, -novelty) %>% 
    group_by(id, group, picture, sleep) %>% 
    summarise(scr = mean(scr_sqrt, na.rm = TRUE),
              arousal = mean(arousal, na.rm = TRUE),
              valence = mean(valence, na.rm = TRUE)) %>% 
    ungroup()

lmer(scr ~ group : picture : sleep + (1 | id), data = sleep_df) %>% 
    summary()

sleep_sum <-
    scr_era %>% 
    filter(session == "emot" & block == 3L |
               session == "recall" & novelty == "old") %>% 
    mutate(sleep = case_when(session == "emot" ~ "pre-sleep",
                             session == "recall" ~ "post-sleep") %>% 
               factor(., c("pre-sleep", "post-sleep"))) %>% 
    select(-session, -block, -novelty) %>% 
    group_by(group, picture, sleep) %>% 
    summarise(scr_mean = mean(scr_sqrt, na.rm = TRUE),
              scr_se = sd(scr_sqrt, na.rm = TRUE)/sqrt(n()),
              arousal_mean = mean(arousal, na.rm = TRUE),
              arousal_se = sd(arousal, na.rm = TRUE)/sqrt(n()),
              valence_mean = mean(valence, na.rm = TRUE),
              valence_se = sd(valence, na.rm = TRUE)/sqrt(n())) %>% 
    ungroup()

sleep_sum %>% 
    ggplot() +
    aes(x = sleep, y = scr_mean, ymin = scr_mean - scr_se, ymax = scr_mean + scr_se, group = picture) +
    geom_point() +
    geom_line(aes(linetype = picture), size = 1.2) +
    geom_errorbar(width = .2) +
    facet_grid(group ~ .)

```

# HDR analysis coming soon
```{r}
```












